---
title: "Machine learning Assignment 2"
author: 'Laxman Thakur - 10393065 -Batch D '
date: "4 August 2019"
output:
  html_document: default
  word_document: default
---

### To clear Global Enviornment

```{r}
rm(list = ls())
```

### To install ggplot2 package

```{r}
install.packages("ggplot2",repos = "http://cran.us.r-project.org")
```

###To load ggplot2 package
```{r}
library(ggplot2)
```

# To install colorspace package
```{r}
install.packages("colorspace",repos = "http://cran.us.r-project.org")
```
### Install H2o package 

```{r}
install.packages("h2o",repos = "http://cran.us.r-project.org")
```

### Load H2o package

```{r}
library(h2o)
```

### To initialize the h2o package
```{r}
h2o.init()
```

### For cluster information

```{r}
h2o.clusterInfo()
```

### To install jpeg package

```{r}
install.packages("jpeg",repos = "http://cran.us.r-project.org")
```

### To load jpeg package

```{r}
library('jpeg')
```

### for Reading Bird image 

```{r}
Bird_image <- readJPEG("Bird Image for CA2.jpg") 
head(Bird_image)
```

###  To show Bird image
```{r}
plot(0:1,0:1,type="n",ann=FALSE,axes=FALSE)
rasterImage(Bird_image,0,0,1,1)
```

### To obtain Bird Image Dimension

```{r}
Bird_image_DM <- dim(Bird_image) 
Bird_image_DM
```


### To assign RGB channels to data frame

```{r}
Bird_imageRGB <-data.frame(
    x=rep(1:Bird_image_DM[2],each=Bird_image_DM[1]),
    y=rep(Bird_image_DM:1,Bird_image_DM[2]),
    R=as.vector(Bird_image[,,1]),
    G= as.vector(Bird_image[,,2]), 
    B= as.vector(Bird_image[,,3])
)

head(Bird_imageRGB)
```

### To use ggplot theme

```{r}
plottheme <-function (){
  theme(     
    panel.background=element_rect(size=3,                                  
                                  colour = "black",                                  
                                  fill = "white"),    
    axis.ticks=element_line(size = 2),    
    panel.grid.major = element_line(colour = "gray80",                                    
                                    linetype = "dotted"),    
    panel.grid.minor = element_line(colour = "gray90",                                    
                                    linetype = "dashed"),
             axis.title.x = element_text(size = rel(1.2),                               
                                         face = "bold"),    
             axis.title.y = element_text(size = rel(1.2),                              
                                         face = "bold"),    
    plot.title = element_text(size = 20, face = "bold",                             
                              vjust = 1.5) ) 
  } 
```


### Plotting the Original Bird image

```{r}
ggplot(data = Bird_imageRGB, aes(x=x, y=y))+  
  geom_point(colour=rgb(Bird_imageRGB[c("R", "G", "B")]))+  
  labs(title = "Original Image: Bird")+ 
  xlab("x")+ 
  ylab("y")+
  plottheme()

```

### colour to Bird image

```{r}
Bird_hex <-as.h2o(Bird_imageRGB)
```
### coefficient of R G B with respect to image for first 6 rows
```{r}
head(Bird_hex)
```


### K means clustering for 2 with h2o

```{r}
kmeans.2 <- h2o.kmeans(training_frame = Bird_hex, k = 2, x = c('R','G','B'))
```
### image centroid coefficient with R G B 
```{r}
kmeans.2@model$centers

```

### image cluster prediction with h2o 
```{r}
kmeans2cluster <- h2o.predict(kmeans.2, Bird_hex[,c('R','G','B')])
```

### coefficient of centroid R G B with respect to k2 cluster for first 6 rows
```{r}

head(as.data.frame(kmeans.2@model$centers[as.vector(kmeans2cluster)+1,]))

```

### To assign RGB channel to data frame of k mean clustering with value 2

```{r}
rgbcluster2 <- as.data.frame(kmeans.2@model$centers[as.vector(kmeans2cluster)+ 1,])[,c('r','g','b')]
kcolors.2 <- rgb(rgbcluster2) 
```

### ggplot of Bird image with k means 2
```{r}
ggplot(data = Bird_imageRGB, aes(x=x, y=y))+  
  geom_point(colour=kcolors.2)+   
  labs(title = paste(2, "Colour of Bird"))+  
  xlab("x")+ 
  ylab("y")+  
  plottheme()

```

### K means clustering for 3 with h2o

```{r}
kmeans.3 <- h2o.kmeans(training_frame = Bird_hex, k = 3, x = c('R','G','B'))
```

### image centroid coefficient with R G B 
```{r}
kmeans.3@model$centers
```

### image cluster prediction with h2o 
```{r}
kmeans3cluster <- h2o.predict(kmeans.3, Bird_hex[,c('R','G','B')])
```

### coefficient of centroid R G B with respect to k3 cluster for first 6 rows
```{r}
head(as.data.frame(kmeans.3@model$centers[as.vector(kmeans3cluster)+1,]))

```

### To assign RGB channel to data frame of k mean clustering with value 3

```{r}
rgbcluster3 <- as.data.frame(kmeans.3@model$centers[as.vector(kmeans3cluster)+ 1,])[,c('r','g','b')] 
kcolors.3 <- rgb(rgbcluster3) 
```

### ggplot of Bird image with k means 3
```{r}
ggplot(data = Bird_imageRGB, aes(x=x, y=y))+   
  geom_point(colour=kcolors.3)+   
  labs(title = paste(3, "Colours of Bird"))+  
  xlab("x")+ 
  ylab("y")+  
  plottheme()

```

### K means clustering for 4 with h2o

```{r}
kmeans.4 <- h2o.kmeans(training_frame = Bird_hex, k = 4, x = c('R','G','B'))
```

### image centroid coefficient with R G B 
```{r}
kmeans.4@model$centers
```

### image cluster prediction with h2o 
```{r}
kmeans4cluster <- h2o.predict(kmeans.4, Bird_hex[,c('R','G','B')])
```

### coefficient of centroid R G B with respect to k4 cluster for first 6 rows
```{r}
head(as.data.frame(kmeans.4@model$centers[as.vector(kmeans4cluster)+1,]))
```

### To assign RGB channel to data frame of k mean clustering with value 4

```{r}
rgbcluster4 <- as.data.frame(kmeans.4@model$centers[as.vector(kmeans4cluster)+ 1,])[,c('r','g','b')] 
kcolors.4 <- rgb(rgbcluster4) 
```

### ggplot of Bird image with k means 4
```{r}
ggplot(data = Bird_imageRGB, aes(x=x, y=y))+   
  geom_point(colour=kcolors.4)+   
  labs(title = paste(4, "Colours of Bird"))+ 
  xlab("x")+ 
  ylab("y")+ 
  plottheme()

```


## K means clustering for 5 with h2o

```{r}
kmeans.5 <- h2o.kmeans(training_frame = Bird_hex, k = 5, x = c('R','G','B'))

```

### image centroid coefficient with R G B 
```{r}
kmeans.5@model$centers
```

### image cluster prediction with h2o 
```{r}
kmeans5cluster <- h2o.predict(kmeans.5, Bird_hex[,c('R','G','B')])
```

### coefficient of centroid R G B with respect to k5 cluster for first 6 rows
```{r}
head(as.data.frame(kmeans.5@model$centers[as.vector(kmeans5cluster)+1,]))

```

### To assign RGB channel to data frame of k mean clustering with value 5

```{r}
rgbcluster5 <- as.data.frame(kmeans.5@model$centers[as.vector(kmeans5cluster)+ 1,])[,c('r','g','b')] 
kcolors.5 <- rgb(rgbcluster5) 

```

### ggplot of Bird image with k means 5
```{r}
ggplot(data = Bird_imageRGB, aes(x=x, y=y))+   
  geom_point(colour=kcolors.5)+   
  labs(title = paste(5, "Colours of Bird"))+  
  xlab("x")+  
  ylab("y")+  
  plottheme()
```


###  sum of squared errors

```{r}
sum_errors <- function(predicted, actuals) { 
  sum ( (predicted-actuals)^2, na.rm=TRUE )
} 
```


## elbow method plot

```{r}
Elbow <- c(sum_errors(rgbcluster2, Bird_imageRGB[,c('R','G','B')]), 
           sum_errors(rgbcluster3, Bird_imageRGB[,c('R','G','B')]),      
           sum_errors(rgbcluster4, Bird_imageRGB[,c('R','G','B')]),         
           sum_errors(rgbcluster5, Bird_imageRGB[,c('R','G','B')])) 

plot(Elbow, type = 'b')

```


### Elbow method with clustering 
```{r}
k <- 1:15 
Elbow <- vector() 
for (i in k){ 
  kmeans <- h2o.kmeans(training_frame = Bird_hex, 
                       k = i, x = c('R','G','B')) 
  kmeanscluster <- h2o.predict(kmeans, Bird_hex[,c('R','G','B')]) 
  rgbcluster <- as.data.frame(kmeans@model$centers[as.vector(kmeanscluster)+1,]) [,c('r','g','b')] 
  Elbow <- append(Elbow, sum_errors(rgbcluster, Bird_imageRGB[,c('R','G','B')]))
}
```

### Graph of sum of squared error with number of clusters

```{r}

plot(Elbow, type='b', xlab = 'Number of Clusters', ylab = 'Sum of Squared Errors')


```

##An elbow chart show the SSE (Sum of Squared Error) after running k-means clustering for k going from 1 to 15. 
##5 seems to be optimal number of clusters.
